---
title: "Files, batches, media and live sessions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Files, batches, media and live sessions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(lm15)
openai <- new_lm("openai", api_key = "sk-example")
```

Every operation on this page has two halves: a pure one that builds the HTTP
request or reads a reply (used below, no network), and the call that sends it.

## Images, audio and documents

Media parts take bytes, a URL, a provider file id, or a local path read when
the request is sent:

```{r}
pdf <- document_part(url = "https://example.org/report.pdf")
req <- request("claude-haiku-4-5", list(message_user(list("Summarise this report.", pdf))))
wire <- build_request(new_lm("anthropic", api_key = "sk-ant-example"), req)
cat(substr(rawToChar(wire$body), 1, 160))
```

A provider with no slot for a kind of media refuses it by name rather than
sending the request without it.

## Files

```{r}
upload <- file_upload_request("notes.txt", bytes_data = charToRaw("Meeting notes"), media_type = "text/plain")
wire <- file_op_build(openai, "upload", upload_request = upload)
wire$method
wire$url
```

`file_upload(openai, upload)` sends it and returns a `file_info()`; use its
`id` in `image_part(file_id = )` and friends.

## Batches

A batch sends many requests at half price, answered within a day.
`batch()` returns a handle: reading it never contacts the provider,
`refresh()` fetches a new snapshot, and only `wait()` polls.

```{r, eval = FALSE}
job <- batch(openai, batch_request(list(
  request("gpt-5-mini", list(message_user("A red fruit?"))),
  request("gpt-5-mini", list(message_user("A green fruit?"))))))
job <- wait(job, poll_every = 60, timeout = 24 * 3600)
for (entry in results(job)) cat(entry$index, response_text(entry$response), "\n")
```

## Generating images, speech and video

```{r}
img <- image_generation_request("gpt-image-1", "A lighthouse at dusk", size = "1024x1024")
generation_build(openai, img)$url
```

`image_generate(openai, img)` returns the images as parts; `speech_generate()`
returns audio. Video is a job: `video_generate()` returns a handle to `wait()` on.

## Live sessions

A live session is a two-way conversation over a WebSocket (OpenAI Realtime,
Gemini Live). The codec is shared with the harness that checks it; here it
builds the frames a session would send first:

```{r}
cfg <- live_config("gpt-realtime", system = "Be brief.", voice = "marin")
frames <- live_setup_frames(openai, cfg)
cat(as_json(frames[[1]]))
```

```{r, eval = FALSE}
session <- live(openai, cfg)
session$send(live_client_text_event("Say hello."))
answer <- result(turn(session))
answer$text
session$close()
```

A turn that exceeds its event or byte budget stops with a
`CollectionLimitError` that keeps what it collected; it never pretends the turn
ended.
