---
title: "Tools, structured output and judgments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tools, structured output and judgments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

As in `vignette("lm15")`, the replies on this page come from a scripted
stand-in for a local Ollama server, written the way Ollama writes them.

```{r}
library(lm15)
chat_reply <- function(message, finish = "stop") list(status = 200L, body = as_json(json_object(
  id = "chatcmpl-1", model = "llama3.2",
  choices = json_array(json_object(index = 0L, finish_reason = finish, message = message)),
  usage = json_object(prompt_tokens = 30L, completion_tokens = 12L))))
```

## Tools

You describe a tool with a JSON Schema. lm15 never derives one from an R
function: the schema is what the model sees, so you write it.

```{r}
weather <- function_tool(
  "weather",
  description = "Current weather in a city.",
  parameters = json_object(
    type = "object",
    properties = json_object(city = json_object(type = "string", description = "City name")),
    required = json_array("city")))
```

The model asks for a call; you run it, send the result back in a tool message,
and ask again. lm15 does not run tools or loop for you.

```{r}
server <- fake_transport(list(
  chat_reply(json_object(role = "assistant", content = NULL, tool_calls = json_array(json_object(
    id = "call_1", type = "function",
    `function` = json_object(name = "weather", arguments = '{"city": "Oslo"}')))), "tool_calls"),
  chat_reply(json_object(role = "assistant", content = "It is 12 degrees and raining in Oslo."))))
router <- new_router(transport = server)

req <- request("ollama:llama3.2", list(message_user("What is the weather in Oslo?")), tools = list(weather))
answer <- complete(router, req)
answer$finish_reason
call <- tool_calls(answer)[[1]]
call$name
call$input$city

result <- "12 degrees C, rain"                      # your code runs the tool here
req$messages <- c(req$messages, list(answer$message, message_tool(call$id, result)))
response_text(complete(router, req))
```

`config(tool_choice = tool_choice(mode = "required"))` forces a call; a
provider that would ignore that instruction is refused before sending.

## Structured output

Ask for JSON that follows a schema with `response_format`:

```{r}
format <- json_object(type = "json_schema", name = "city", strict = TRUE, schema = json_object(
  type = "object",
  properties = json_object(name = json_object(type = "string"), population = json_object(type = "integer")),
  required = json_array("name", "population"),
  additionalProperties = FALSE))
server <- fake_transport(list(chat_reply(json_object(role = "assistant", content = '{"name": "Oslo", "population": 717710}'))))
answer <- complete(new_router(transport = server),
                   request("ollama:llama3.2", list(message_user("Largest city in Norway?")),
                           config = config(response_format = format)))
parse_json(answer)$population
```

Numbers keep their exact spelling; large integers are never rounded.

## Judgments

A judgment is a question whose answer is one of the keys you declare. Some
providers can also say how likely each key was. `judgments()` writes the
schema from `yes_no()`, `choice()` and `score()`:

```{r}
review <- judgments(
  quality = score("How good is the wine, according to the note?",
                  c(poor = "Faulty or unpleasant", fine = "Sound, simple", great = "Complex, long")),
  style = choice("Which style dominates?", c(fruit = "Fruit-forward", oak = "Oak-driven", other = NA)),
  ageing = yes_no("Does the note say it will improve with age?"))
note <- message_user("Ripe blackberry and cassis, toasty oak; will reward a decade in the cellar.")
```

`config(probabilities = )` says what to do when a provider can only give the
pick: `"if_available"` answers with the pick and records that no distribution
was measured; `"required"` refuses before sending.

```{r}
server <- fake_transport(list(chat_reply(json_object(role = "assistant",
  content = '{"quality": 2, "style": "oak", "ageing": true}'))))
answer <- complete(new_router(transport = server),
  request("ollama:llama3.2", list(note), config = config(response_format = review, probabilities = "if_available")))
response_data(answer)$style
response_probabilities(answer)
answer$adaptations[[1]]$field
```

TypeSafe's `jev-` models measure a distribution for every judgment:

```{r}
jev <- fake_transport(list(list(status = 200L, body = as_json(json_object(model = "jev-latest",
  answers = json_object(
    quality = json_object(type = "score", probabilities = json_object("0" = 0.02, "1" = 0.18, "2" = 0.80)),
    style = json_object(type = "choice", choice = "oak", probabilities = json_object(fruit = 0.3, oak = 0.65, other = 0.05)),
    ageing = json_object(type = "noul", noul = 0.93)))))))
router <- new_router(api_keys = list(typesafe = "ts-example"), transport = jev)
answer <- complete(router, request("jev-latest", list(note), config = config(response_format = review, probabilities = "required")))
response_data(answer)$quality
response_probabilities(answer)$ageing$true
response_expected(answer, "quality")
response_method(answer)
```

A vLLM server that scores named tokens (`new_lm("vllm")`) measures the same
distribution from the model's own token probabilities, in one batched call.
No number the model wrote in its text is ever read as a probability.

## Structured input

A data part carries any JSON value. On a provider that reads only text it goes
as compact JSON, with nothing added:

```{r}
row <- data_part(json_object(note = "Ripe cassis", price_eur = 48L))
wire <- build_request(new_lm("ollama"), request("llama3.2", list(message_user(row))))
cat(rawToChar(wire$body))
```
