---
title: "Getting started with lm15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with lm15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

lm15 gives you one way to talk to many language model providers. You build a
request, ask a client to send it, and keep the whole response. Nothing is sent,
retried, shortened or remembered behind your back.

## A stand-in server for this page

Every example on this page really runs: requests are built for the provider's
wire, sent, and the replies parsed. The replies come from a scripted stand-in
for a local [Ollama](https://ollama.com) server, so this page needs no key and
no network. `chat_reply()` writes a reply the way Ollama would.

```{r}
library(lm15)

chat_reply <- function(text, finish = "stop") list(status = 200L, body = as_json(json_object(
  id = "chatcmpl-1", model = "llama3.2",
  choices = json_array(json_object(index = 0L, finish_reason = finish,
    message = json_object(role = "assistant", content = text))),
  usage = json_object(prompt_tokens = 21L, completion_tokens = 9L))))

server <- fake_transport(list(
  chat_reply("A list column holds a whole list in each row of a data frame."),
  chat_reply("Use tidyr::unnest() to turn each element into its own row.")))
router <- new_router(transport = server)
```

In your own code, `new_router()` needs no `transport`: it sends over the
network, and reads keys from the usual environment variables
(`OPENAI_API_KEY`, `ANTHROPIC_API_KEY`, ...) when a request is sent.

## One request, one response

```{r}
req <- request(
  "ollama:llama3.2",
  list(message_user("What is a list column?")),
  system = "Answer in one sentence.",
  config = config(max_tokens = 100L, temperature = 0.2)
)
answer <- complete(router, req)
response_text(answer)
answer$finish_reason
answer$usage$input_tokens
```

The model string is `provider:model`. A router also routes well-known bare
names (`gpt-5-mini` to OpenAI, `claude-haiku-4-5` to Anthropic, ...):

```{r}
resolve(router, "claude-haiku-4-5")
```

## A conversation is a list you keep

lm15 keeps no hidden history. To continue, append the assistant's message and
yours, and send the new request:

```{r}
req$messages <- c(req$messages, list(answer$message, message_user("How do I expand one?")))
response_text(complete(router, req))
```

Changing a field of a request returns a new request, and the change is
checked: a tool result in a user message, for example, is refused on the spot.

## What was actually sent

`recorded_requests()` shows what the stand-in received; with a real provider,
`build_request()` shows the same thing without sending it.

```{r}
sent <- recorded_requests(server)
length(sent)
cat(rawToChar(sent[[1]]$body))
```

## When a provider cannot do what you asked

Providers differ. When a wire has no slot for a setting, lm15 adapts and says
so on the response, or refuses before sending. It never drops a setting
without a record. `plan()` shows what would happen, without sending anything:

```{r}
req <- request("openai:gpt-5-mini", list(message_user("Hi")),
               config = config(top_k = 40L, stop = list("END")))
for (a in plan(new_router(), req)) cat(a$field, "->", a$action, "\n")
```

`new_router(adaptations = "refuse")` turns every such adaptation into an
error instead.

## Errors

Errors are conditions with the same classes in every lm15 language:
`RateLimitError`, `AuthError`, `ContextLengthError`, `UnsupportedModelError`,
and so on, all inheriting from `LM15Error`. lm15 never retries; `retryable()`
tells you whether retrying could help.

```{r}
busy <- fake_transport(list(list(status = 429L, headers = list("retry-after" = "2"),
  body = '{"error": {"message": "Too many requests", "type": "rate_limit_exceeded"}}')))
err <- tryCatch(complete(new_router(transport = busy), request("ollama:llama3.2", list(message_user("hi")))),
                LM15Error = identity)
class(err)[1]
retryable(err)
err$retry_after
```

## Streaming

`stream()` calls your function with each event as it arrives, and returns the
assembled response at the end.

```{r}
sse <- paste0(
  'data: {"choices":[{"index":0,"delta":{"content":"Rows "}}]}\n\n',
  'data: {"choices":[{"index":0,"delta":{"content":"of lists."}}]}\n\n',
  'data: {"choices":[{"index":0,"delta":{},"finish_reason":"stop"}]}\n\n',
  'data: [DONE]\n\n')
streaming <- new_router(transport = fake_transport(list(list(status = 200L, chunks = list(sse)))))
final <- stream(streaming, request("ollama:llama3.2", list(message_user("Define a list column."))),
                function(event) if (event$type == "delta") cat(event$delta$text))
final$finish_reason
```

## Where to go next

* `vignette("conversations-and-tools")`: tools, structured output and judgments.
* `vignette("signing-in")`: keys, subscriptions, cloud identities, and what lm15
  will and will not use.
* `vignette("beyond-chat")`: files, batches, media generation and live sessions.
