---
title: "Keys, subscriptions and cloud identities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Keys, subscriptions and cloud identities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(lm15)
```

## Keys from the environment

With no other configuration, a router reads a provider's usual environment
variable when a request is sent: `OPENAI_API_KEY`, `ANTHROPIC_API_KEY`,
`GEMINI_API_KEY`, and so on. Nothing is read when the router is created.

`explain_auth()` shows which credential a request would use, and why, without
printing a secret or contacting anything:

```{r}
explain_auth("anthropic", env = c(ANTHROPIC_API_KEY = "sk-ant-example"))
```

Explicit credentials win over everything:

```{r}
router <- new_router(api_keys = list(anthropic = "sk-ant-from-a-vault"))
```

A credential may also be a function that returns one; it is called for every
request, so a rotating token is never cached by lm15.

## Subscriptions

Some providers let you use a subscription instead of a paid key: xAI (SuperGrok
or X Premium), Claude, ChatGPT, GitHub Copilot, Kimi Code, Meta. Sign in once:

```{r, eval = FALSE}
login("xai")            # shows a code to approve in your browser
router <- new_router(auth = local_auth())
```

The login is saved in a private file that lm15 in Python, TypeScript, Rust and
Go read too (`~/.config/lm15/credentials.json` unless `LM15_CREDENTIALS_PATH`
says otherwise). It is renewed when due, under a lock the other languages
respect, so a refresh token is never spent twice.

A failed or signed-out subscription is **not** replaced by a paid key on its
own: you would be charged without having chosen to be. Pass the key explicitly
if that is what you want. The same rules, shown with a scope that lives only
in this R session:

```{r}
auth <- memory_auth()
key <- set_api_key("openai", "sk-example", auth = auth)
status("openai", auth = auth)$usability
logout("openai", auth = auth)$forgot
env <- c(OPENAI_API_KEY = "sk-ambient")
report <- explain_auth("openai", env = env, auth = auth)
report$configured
vapply(report$steps, function(s) paste(s$kind, s$state), "")
```

Other things you can save: a key's environment variable *name* (read at
request time, never copied), a local server, or a named cloud identity:

```{r}
configure("gemini", "env", answers = list(name = "GEMINI_API_KEY"), auth = auth)
configure("ollama", "local", answers = list(base_url = "http://127.0.0.1:11434/v1"), auth = auth)
vapply(connections(auth = auth), function(c) c$label, "")
```

## connect()

At a console, `connect()` asks which provider, signs in or saves a key if
needed, lists your account's models, and returns a client bound to that
connection and model. It never sends a prompt.

```{r, eval = FALSE}
client <- connect()
response_text(complete(client, "Hello"))
```

`bind_model()` does the same without questions:

```{r}
auth <- memory_auth()
set_api_key("openai", "sk-example", auth = auth)
client <- bind_model("openai", "gpt-5-mini", auth = auth)
client$request("Hello")$model
```

A bound client follows its connection's renewals and nothing else: if the
connection is replaced or signed out, it stops rather than switch who pays.

## Cloud identities

Azure, AWS and Google Cloud providers walk the cloud's own chain (environment,
profile, CLI login, the machine's identity) and say which one answered.
`explain_auth()` also reports where each host setting came from:

```{r}
explain_auth("vertex", env = c(GOOGLE_CLOUD_PROJECT = "my-project", NO_GCE_CHECK = "1", HOME = tempdir()))
```

To use exactly one identity and nothing else, name it: `platform` (the
machine's own), `workload`, `environment` or `cli`.

```{r, eval = FALSE}
router <- new_router(credentials = list(bedrock_chat = "cli"),
                     settings = list(bedrock_chat = list(region = "eu-west-3")))
```
